A simplified Plackett-Luce likelihood function for rank orders

### 2024-03-04 inst/ejs-sample.tex
rewording

One issue encountered in educational research is that of
confidentiality of examination scores, and individual students'
results are generally regarded as sensitive and private information.
Nonetheless, it is common for students to possess two pertinent and
informative observations: the total enrollment in a specific class and
their individual ranking within that class; table~\ref{educationtable}
shows an apocryphal dataset.  What could a student infer from such a
dataset?

@@

One natural question would be to wonder whether the student was bettar
at pure or applied mathematics and, if so, to quantify the difference.
Figure~\ref{plotpureandapplied} shows contours of support [again
  normalised so the support of the evaluate is zero] for the dataset,
as a function of pure strength on the horizontal axis and applied
strength on the vertical axis.  A null of equal strengths appears as a
diagonal line.  We see that the maximum likelihood estimate is
$\hat{p}_\mathrm{pure}=0.567$, $\hat{p}_\mathrm{applied}=0.887$.
However, a constrained optimization shows that the maximum support
along the null diagonal is about $-2.43$.  Edwards's two units of
support criterion is met, and we are justified in concluding that the
student is indeed more competent at applied than pure mathematics.
Alternatively, we may observe that $2\mathcal{S}=4.86$ is in the tail
region of its asymptotic $\chi^2_1$ distribution, corresponding to a
$p$-value of about $0.028$.

>>>>

One issue encountered in educational research is that of
confidentiality of examination scores, and individual results are
generally regarded as sensitive and private information.
Nevertheless, it is common a student to possess some informative
observations: the total enrollment in a specific class and their
individual ranking within that class;  table~\ref{educationtable}
shows an apocryphal dataset.   What could a student infer from such
data?

@@

One natural question would be to wonder whether the student was better
at pure or applied mathematics and, if so, to quantify the difference.
Figure~\ref{plotpureandapplied} shows contours of support [again
  normalised so the support of the evaluate is zero] for the dataset,
as a function of pure strength on the horizontal axis and applied
strength on the vertical axis.  A null of equal strengths appears as a
diagonal line.  We see that the maximum likelihood estimate is
$\hat{p}_\mathrm{pure}=0.567$, $\hat{p}_\mathrm{applied}=0.887$.
However, a constrained optimization shows that the maximum support
along the null diagonal is about $-2.43$.  Edwards's two units of
support criterion is met, and we are justified in concluding that the
student is indeed more competent at applied than pure mathematics.
Alternatively, we may observe that $2\mathcal{S}=4.86$ is in the tail
region of its asymptotic $\chi^2_1$ distribution, corresponding to a
$p$-value of about $0.028$.
